Elaboration with First-Class Implicit Function Types

### 2020-02-25 paper.tex
stuff

\begin{example}
We illustrate now the action of constancy constraints, using the example from
Section \ref{sec:superfluous}. Again, assume $Bool : \U$ and $true : Bool$.
\begin{alignat*}{3}
& \scriptstyle{0}\qquad\qquad && \einfert{\slet\,x : \_ = true\,\sin\,x}{\true}
      \hspace{21em}\\
& \scriptstyle{1} && \quad \echeckt{\_}{\U}\\
& \scriptstyle{2} && \quad = \alpha_0\\
& \scriptstyle{3} && \quad \echeckt{true}{\alpha_0}\\
& \scriptstyle{4} && \qquad \mathsf{freshMeta}\,\Tel\\
& \scriptstyle{5} && \qquad = \alpha_1\\
& \scriptstyle{6} && \qquad \einfert{true}{\true}\\
& \scriptstyle{7} && \qquad = true : Bool\\
& \scriptstyle{8} && \qquad \mathsf{newConstancy}_{\emptycon,\,x : \Rec\,\alpha_1}\,Bool\\
& \scriptstyle{9} && \qquad\quad \unify\,\alpha_1\,\epsilon\\
& \scriptstyle{10} && \qquad \unify\,\alpha_0\,Bool\\
& \scriptstyle{11} && \quad = true\\
& \scriptstyle{12} && \quad \einfert{x}{\true}\\
& \scriptstyle{13} && \quad = true : Bool\\
& \scriptstyle{14} && = (\slet\,x : Bool = true\,\sin\,x) : Bool
\end{alignat*}
First we create a fresh meta $\alpha_0$ for the hole on line $\scriptstyle{2}$,
then we check $\true$ with it. Since we are checking with a meta-headed type, we
create a fresh telescope meta $\alpha_1$, then infer $Bool$ for $true$. On line
$\scriptstyle{8}$ we create a constancy constraint, and immediately try to solve
it, in accordance with the algorithmic implementation in Section
\ref{sec:algorithmic}. Since $Bool$ does not depend on the domain, we solve the
constraint and thereby solve $\alpha_1$ as $\epsilon$. Hence, we simply return
$true$ on line $\scriptstyle{11}$, since $\lambda\,\{x : \ol{\epsilon}\}.\,true$
computes to that, and elaboration succeeds with the expected output.
\end{example}

@@

\begin{figure}
\begin{tabular}{|lll|}
\hline
A & \multicolumn{2}{l|}{\textsc{polymorphic instantiation}}\\
\hline
A1 & $\lambda\,x\,y.\,y$ & \yesst\\
A2 & $choose\,id$ & \yesst \\
A3 & $choose\,nil\,ids$ & \yes \\
A4 & $\lambda\,(x : \{A : \U\}\to A \to A).\,x\,x$ & \yesst\\
A5 & $id\,auto$ & \yes\\
A6 & $id\,auto'$ & \yesst\\
A7 & $choose\,id\,auto$ & \yes\\
A8 & $choose\,id\,auto'$ & \no\\
A9 & $\lambda\,(f : \{A : \U\}\to(A\to A)\to List\,A \to A).\,f\,(choose\,id)\,ids$ & \yesst\\
A10 & $poly\,id$ & \yes\\
A11 & $poly\,(\lambda\,x.\,x)$ & \yes\\
\hline
B & \multicolumn{2}{l|}{\textsc{inference of polymorphic arguments}}\\
\hline
B1 & $\lambda\,f.\,pair\,(f\,zero)\,(f\,true)$ & \no\\
B2 & $\lambda\,xs.\,poly\,(head\,xs)$ & \no\\
\hline
C & \multicolumn{2}{l|}{\textsc{functions on polymorphic lists}}\\
\hline
C1  & $length\,ids$ & \yes \\
C2  & $tail\,ids$ & \yes \\
C3  & $head\,ids$ & \yesst \\
C4  & $single\,id$ & \yesst \\
C5  & $cons\,id\,ids$ & \yes \\
C6  & $cons\,(\lambda\,x.\,x)\,ids$ & \yes \\
C7  & $append\,(single\,inc)\,(single\,id)$ & \yes \\
C8  & $\lambda\,(g:\{A:\U\}\to List\,A \to List\,A \to A).\,g\,(single\,id)\,ids$ & \yesst \\
C9  & $map\,poly\,(single\,id)$ & \yes \\
C10 & $map\,head\,(single\,ids)$ & \yes \\
\hline
D & \multicolumn{2}{l|}{\textsc{application functions}}\\
\hline
D1 & $app\,poly\,id$        & \yes \\
D2 & $revapp\,id\,poly$     & \yes \\
D3 & $runST\,argST$         & \yes \\
D4 & $app\,runST\,argST$    & \yes \\
D5 & $revapp\,argST\,runST$ & \yes \\
\hline
\end{tabular}

@@

%% λ (g : {A} → List A → List A → A). g (single id) ids

>>>>

%% \begin{alignat*}{3}
%% & \scriptstyle{0}\qquad\qquad && \einfert{\slet\,x : \_ = true\,\sin\,x}{\true}
%%       \hspace{21em}\\
%% & \scriptstyle{1} && \quad \echeckt{\_}{\U}\\
%% & \scriptstyle{2} && \quad = \alpha_0\\
%% & \scriptstyle{3} && \quad \echeckt{true}{\alpha_0}\\
%% & \scriptstyle{4} && \qquad \mathsf{freshMeta}\,\Tel\\
%% & \scriptstyle{5} && \qquad = \alpha_1\\
%% & \scriptstyle{6} && \qquad \einfert{true}{\true}\\
%% & \scriptstyle{7} && \qquad = true : Bool\\
%% & \scriptstyle{8} && \qquad \mathsf{newConstancy}_{\emptycon,\,x : \Rec\,\alpha_1}\,Bool\\
%% & \scriptstyle{9} && \qquad\quad \unify\,\alpha_1\,\epsilon\\
%% & \scriptstyle{10} && \qquad \unify\,\alpha_0\,Bool\\
%% & \scriptstyle{11} && \quad = true\\
%% & \scriptstyle{12} && \quad \einfert{x}{\true}\\
%% & \scriptstyle{13} && \quad = true : Bool\\
%% & \scriptstyle{14} && = (\slet\,x : Bool = true\,\sin\,x) : Bool
%% \end{alignat*}

\begin{example}
We illustrate now the action of constancy constraints, using the example from
Section \ref{sec:superfluous}. Again, assume $Bool : \U$ and $true : Bool$.
\begin{alignat*}{3}
& \scriptstyle{0}\qquad\qquad && \einfert{\slet\,x : \_ = true\,\sin\,x}{\true}
      \hspace{21em}\\
& \scriptstyle{1} && \quad \echeckt{\_}{\U} = \alpha_0\\
& \scriptstyle{3} && \quad \echeckt{true}{\alpha_0}\\
& \scriptstyle{4} && \qquad \mathsf{freshMeta}\,\Tel = \alpha_1\\
& \scriptstyle{6} && \qquad \einfert{true}{\true} = true : Bool\\
& \scriptstyle{8} && \qquad \mathsf{newConstancy}_{\emptycon,\,x : \Rec\,\alpha_1}\,Bool\\
& \scriptstyle{9} && \qquad\quad \unify\,\alpha_1\,\epsilon\\
& \scriptstyle{10} && \qquad \unify\,\alpha_0\,Bool\\
& \scriptstyle{11} && \quad = true\\
& \scriptstyle{12} && \quad \einfert{x}{\true} = true : Bool\\
& \scriptstyle{14} && = (\slet\,x : Bool = true\,\sin\,x) : Bool
\end{alignat*}
First we create a fresh meta $\alpha_0$ for the hole on line $\scriptstyle{2}$,
then we check $\true$ with it. Since we are checking with a meta-headed type, we
create a fresh telescope meta $\alpha_1$, then infer $Bool$ for $true$. On line
$\scriptstyle{8}$ we create a constancy constraint, and immediately try to solve
it, in accordance with the algorithmic implementation in Section
\ref{sec:algorithmic}. Since $Bool$ does not depend on the domain, we solve the
constraint and thereby solve $\alpha_1$ as $\epsilon$. Hence, we simply return
$true$ on line $\scriptstyle{11}$, since $\lambda\,\{x : \ol{\epsilon}\}.\,true$
computes to that, and elaboration succeeds with the expected output.
\end{example}

@@

{\small

\begin{figure}
\begin{tabular}{|lll|}
\hline
A & \multicolumn{2}{l|}{\textsc{polymorphic instantiation}}\\
\hline
A1 & $\lambda\,x\,y.\,y$ & \yesst\\
A2 & $choose\,id$ & \yesst \\
A3 & $choose\,nil\,ids$ & \yes \\
A4 & $\lambda\,(x : \{A : \U\}\to A \to A).\,x\,x$ & \yesst\\
A5 & $id\,auto$ & \yes\\
A6 & $id\,auto'$ & \yesst\\
A7 & $choose\,id\,auto$ & \yes\\
A8 & $choose\,id\,auto'$ & \no\\
A9 & $\lambda\,(f : \{A : \U\}\to(A\to A)\to List\,A \to A).\,f\,(choose\,id)\,ids$ & \yesst\\
A10 & $poly\,id$ & \yes\\
A11 & $poly\,(\lambda\,x.\,x)$ & \yes\\
\hline
B & \multicolumn{2}{l|}{\textsc{inference of polymorphic arguments}}\\
\hline
B1 & $\lambda\,f.\,pair\,(f\,zero)\,(f\,true)$ & \no\\
B2 & $\lambda\,xs.\,poly\,(head\,xs)$ & \no\\
\hline
C & \multicolumn{2}{l|}{\textsc{functions on polymorphic lists}}\\
\hline
C1  & $length\,ids$ & \yes \\
C2  & $tail\,ids$ & \yes \\
C3  & $head\,ids$ & \yesst \\
C4  & $single\,id$ & \yesst \\
C5  & $cons\,id\,ids$ & \yes \\
C6  & $cons\,(\lambda\,x.\,x)\,ids$ & \yes \\
C7  & $append\,(single\,inc)\,(single\,id)$ & \yes \\
C8  & $\lambda\,(g:\{A:\U\}\to List\,A \to List\,A \to A).\,g\,(single\,id)\,ids$ & \yesst \\
C9  & $map\,poly\,(single\,id)$ & \yes \\
C10 & $map\,head\,(single\,ids)$ & \yes \\
\hline
D & \multicolumn{2}{l|}{\textsc{application functions}}\\
\hline
D1 & $app\,poly\,id$        & \yes \\
D2 & $revapp\,id\,poly$     & \yes \\
D3 & $runST\,argST$         & \yes \\
D4 & $app\,runST\,argST$    & \yes \\
D5 & $revapp\,argST\,runST$ & \yes \\
\hline
E & \multicolumn{2}{l|}{\textsc{$\eta$-expansion}}\\
\hline
& \multicolumn{2}{l|}{assuming $k : \{A : \U\}\to A \to List\,A \to A$,} \\
& \multicolumn{2}{l|}{$h : Int \to \{A:\U\}\to A \to A$}\\
& \multicolumn{2}{l|}{and $lst : List\,(\{A:\U\} \to Int \to A \to A)$} \\
E1 & $k\,h\,lst$ & \no \\
E2 & $k\,(\lambda\,x.\,h\,x)\,lst$ & \yesst \\
E3 & $\lambda\,(r : (\{A :\U\}\to A \to \{B : \U\}\to B \to B) \to Int).\,r\,(\lambda\,x\,y.\,y)$ & \yes \\
\hline
\multicolumn{3}{l}{\text{{\footnotesize
``Yes*'' means that our system can infer a type for the
expression, but the lack of}}}\\
\multicolumn{3}{l}{\text{{\footnotesize
let-generalization causes unsolved metas in the type.}}}
\end{tabular}

@@

}

%% (h : Int → {A} → A → A)(k : {A} → A → List A → A)(lst : List ({A} → Int → A → A))
